Gravitational wave sensitivity curves

### 2014-06-30 Typesofsource.tex
Burst sources

\cjm{This section is currently much shorter than both \ref{sec:insp} and \ref{sec:stoch}, Rob said he was going to add more detail on searching for unmodelled phenomenon using time-frequency plots.}

A signal is burst-like if it's duration at a detectable amplitude is only a few times the wave period. If this is the case then the signal does not have time to accumulate SNR in each frequency band in the way inspirals do. If the source generates ${\cal{N}}_{\textrm{cycles}}$ of radiation then the relation between $h_{c}(f)$ and $h$ is simply
\begin{equation}\label{eq:simple} h_{c}(f)=h\sqrt{{\cal{N}}_{\textrm{cycles}}} \; . \end{equation}

>>>>

Some sources of GWs can produce signals with large amplitudes, greater than the detector noise. The duration of such a signal is typically very short, of the order of a few wave periods, and so there is not time to accumulate SNR in each frequency band in the way inspirals do. As a consequence, waveform models are not required to produce a detection; we simply rely on the "excess power" produced by these so-called burst sources. Typically, we may be looking for signals from core collapse supernovae \citep{Ott2009}, the late stages of merging compact binaries, or more generally, signals from any unexpected or poorly modelled sources.

Burst searches are often carried out using time-frequency techniques. The data stream from a detector is temporally split into segments, the length of which can be tuned to give greater sensitivity to particular sources. Each segment is then transformed into the frequency domain, whitened and normalised to the noise spectrum to produce a time-frequency plot. Potential GW signals are identified by searching for clusters of pixels that contain an excess of power (see e.g. \cite{Bursts}).

The presence of excess power across a number of pixels eliminates modelled noise sources, but such a cluster may also be caused by atypical noise within a detector. We can improve our confidence of a GW signal by making use of information obtained from other GW detectors. Signals across a network of detectors should have compatible arrival times (given the sky direction) as well as consistent amplitudes, frequencies and shapes of the waveform. Different pipelines are currently in use that analyse the signal consistency in different ways: both coincidence searches \citep{Chatterji2004} and fully coherent methods \citep{Klimenko2008} are used.

An important aspect of burst search algorithms is to accurately estimate the noise properties within each time segment. To this extent, null data streams can be constructed that should be insensitive to real GW signals. In order to estimate the false alarm rate, the data from different detectors can be shifted in time to remove any genuine coincident GW signals. These time-shifts are then analysed to simulate the potential occurence of coincident noise events. The algorithms are tuned using time-shifted data to ensure there is no bias in the final search.

As discussed in \ref{sec:insp}, the expected relation between $h_{c}(f)$ and a typical waveform $\tilde{h}(f)$ is simply
\begin{equation}\label{eq:simple} h_{c}(f)=\tilde{h}(f)\sqrt{{\cal{N}}_{\textrm{cycles}}} \; , \end{equation}
where ${\cal{N}}_{\textrm{cycles}}$ is the number of cycles of radiation generated by the source. Here, ${\cal{N}}_{\textrm{cycles}}$ is of order unity and $\tilde{h}(f)$ is strongly peaked around some central frequency.
